---
title: 'DATA 607 Extra Credit'
subtitle: 'Window Functions'
author: "Kristin Lussi"
date: "2023-9-26"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this presentation, I will demonstrate time series analysis on a dataset which provides daily climate data from January 1, 2013 to April 24, 2017 in the city of Delhi, India. The variables measured are mean temperature (celsius), humidity (g.m^-3), wind speed (kmph), and mean pressure (atm). 

```{r, message = FALSE, warning = FALSE}
library(readr)
library(stringr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(gt)

urlfile <- "https://raw.githubusercontent.com/kristinlussi/607ExtraCredit3/main/DailyDelhiClimateTrain.csv"

delhiClimate <- read_csv(url(urlfile), show_col_types = FALSE)
```

Here is a glimpse of the data:
```{r}
head(delhiClimate) 
```
## Yearly Averages

Here, we will calculate the yearly averages for each year (2013 - 2016).

```{r}
yearly_average <- delhiClimate %>%
  mutate(year = year(delhiClimate$date)) %>%
  group_by(year) %>%
  summarize(
    meanTemp = mean(meantemp, na.rm = TRUE),
    humidity = mean(humidity, na.rm = TRUE),
    windSpeed = mean(wind_speed, na.rm = TRUE),
    meanPressure = mean(meanpressure, na.rm = TRUE)
  ) 

# filter out year 2017
yearly_average <- yearly_average %>%
  filter(year != 2017)
```

```{r, include = FALSE}
yearly_average_tbl <- gt(yearly_average) |>
  tab_header(
    title = md("**Yearly Averages for Years 2013 to 2016**"),
    subtitle = "Recorded in Delhi, India"
  ) |>
  cols_label(
    year = md("**Year**"),
    meanTemp = md("**Mean Temp**"),
    humidity = md("**Humidity**"),
    windSpeed = md("**Wind Speed**"),
    meanPressure = md("**Mean Pressure**")
  )
```

```{r}
yearly_average_tbl
```

## 30 Day Moving Average

Here, I calculated the 30-day moving averages for each variable using filter function from the stats package. I used 30 days here instead of 6 because we are looking at a long stretch of data (2013-2016). It is easier to see long-term trends when using a longer moving average. 

```{r}
# calculate 30-day moving average for mean temperature column
delhiClimate$temp_ma <- stats::filter(delhiClimate$meantemp, filter = rep(1/30, 30), sides = 2)

# calculate 30-day moving average for humidity column
delhiClimate$humidity_ma <- stats::filter(delhiClimate$humidity, filter = rep(1/30, 30), sides = 2)

# calculate 30-day moving average for wind speed column
delhiClimate$windspeed_ma <- stats::filter(delhiClimate$wind_speed, filter = rep(1/30, 30), sides = 2)

# calculate 30-day moving average for mean pressure column 
delhiClimate$pressure_ma <- stats::filter(delhiClimate$meanpressure, filter = rep(1/30, 30), sides = 2)
```

## Time Series Plot for Mean Temperature

```{r, warning = FALSE, message = FALSE}
ggplot(data = delhiClimate) +
  geom_line(aes(x = date, y = temp_ma, color = factor(year(date)), group = year(date)), 
            na.rm = TRUE, linewidth = 1) + 
  labs(x = "Year", y = "Mean Temperature (°C)", title = "30 Day Moving Average of Mean Temperature") +
  theme_minimal() +
  theme(legend.position = "none")
```

As you can see from the plot, there is a similar trend from year to year, with higher temperatures in the summer months and lower temperatures in the winter months. 

## Time Series Plot for Humidity
```{r, warning = FALSE, message = FALSE}
ggplot(data = delhiClimate) +
  geom_line(aes(x = date, y = humidity_ma, color = factor(year(date)), group = year(date)), 
            na.rm = TRUE, linewidth = 1) +
  labs(x = "Year", y = "Humidity", title = "30 Day Moving Average of Humidity") +
  theme_minimal() +
  theme(legend.position = "none")
```

As you can see from the above plot, there is a slight trend in humidity levels from year to year. The humidity is lower in the spring and higher in the summer and winter months. 

## Time Series Plot for Wind Speed
```{r, warning = FALSE, message = FALSE}
ggplot(data = delhiClimate) +
  geom_line(aes(x = date, y = windspeed_ma, color = factor(year(date)), group = year(date)), 
            na.rm = TRUE, linewidth = 1) +  
  labs(x = "Year", y = "Wind Speed (kmph)", title = "30 Day Moving Average of Wind Speed") +
  theme_minimal() +
  theme(legend.position = "none")
```

From the above plot, we can see that there is a slight trend in wind speed from year-to-year. Wind speed tends to be highest in the middle of the year (June).

## Time Series Plot for Mean Pressure
```{r, warning = FALSE, message = FALSE}
ggplot(data = delhiClimate) +
  geom_line(aes(x = date, y = pressure_ma, color = factor(year(date)), group = year(date)), 
            na.rm = TRUE, linewidth = 1) +  
  labs(x = "Year", y = "Mean Pressure (atm)", title = "30 Day Moving Average of Mean Pressure") +
  theme_minimal() +
  theme(legend.position = "none")
```

As you can see, it seems that there was a trend in the years 2013-2015, but in 2016 there was an unsual spike and decline in mean pressure levels. 

## Conclusion

From the analysis, we can conclude that there is a\ trend from year-to-year in regards to mean temperature, humidity, and wind speed. However, we cannot conclude that there is a trend from year-to-year in regards to mean pressure, since there was an unusual spike and decline in 2016 which we are not able to explain with the data provided. 

## Sources

### Kaggle Dataset:

Title: Daily Climate Time Series Data

Author/Uploader: Sumanthvrao

URL: https://www.kaggle.com/datasets/sumanthvrao/daily-climate-time-series-data

### TradeStation Help Page:
Title: What is a Trend?

Publisher/Website: TradeStation

URL: https://help.tradestation.com/10_00/eng/tradestationhelp/data/what_trend.html 

### Fidelity
Title: Moving Average Trading Signal - Stocks

Publisher/Website: Fidelity

URL: https://www.fidelity.com/viewpoints/active-investor/moving-averages#:~:text=A%20longer%20moving%20average%2C%20such,to%20assess%20short%2Dterm%20patterns. 

